---
title: "p8105_hw1_scl2160"
author: "Sam Loewen"
date: "9/18/2019"
output: github_document
---

## Problem 1

First I am creating a dataframe that meets these requirements:

*   a random sample of size 8 from a standard Normal distribution
*   a logical vector indicating whether elements of the sample are greater than 0
*   a character vector of length 8
*   a factor vector of length 8, with 3 different factor “levels”

```{r create_df}
library(tidyverse)
set.seed(3471)
df_p1 = tibble(
  norm_samp = rnorm(8),
  log_vec = norm_samp > 0,
  char_vec = c("dog","cat", "puppy", "kitten", "eggs", "sandwich", "beer", "pencil"),
  factor_vec = factor(c("pug","lab","chow","chow","lab","chow","pug","lab"))
)
```

When I take the mean of each variable in my dataframe I can only get the mean of 'norm_samp' and 'log_vec', as they are the only numeric variables.

```{r means}
mean(pull(df_p1, norm_samp))
mean(pull(df_p1, log_vec))
mean(pull(df_p1, char_vec))
mean(pull(df_p1, factor_vec))
```

Next I will write a code chunk that applies the as.numeric function to the logical, character, and factor variables.

```{r as_numeric}
echo = FALSE
as.numeric(pull(df_p1,log_vec))
as.numeric(pull(df_p1,char_vec))
as.numeric(pull(df_p1,factor_vec))
```

My logic vector and factor vector were able to be converted to numeric, but my character variable was not.

I will now: 


*   convert the logical vector to numeric, and multiply the random sample by the result
*   convert the logical vector to a factor, and multiply the random sample by the result
*   convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r var_change}
as.numeric(pull(df_p1, log_vec))*(pull(df_p1, norm_samp))
as.factor(pull(df_p1, log_vec))*(pull(df_p1, norm_samp))
as.numeric((as.factor(pull(df_p1, log_vec))))*(pull(df_p1, norm_samp))
```


## Problem 2

First I will create a data frame comprised of:

*   x: a random sample of size 500 from a standard Normal distribution
*   y: a random sample of size 500 from a standard Normal distribution
*   A logical vector indicating whether x + y > 1
*   A numeric vector created by coercing the above logical vector
*   A factor vector created by coercing the above logical vector

```{r x_y}
library(tidyverse)
set.seed(2345)
df_p2 = tibble(
  x = rnorm(500),
  y = rnorm(500),
  log_vec_2 = x + y > 1,
  num_vec_2 = as.numeric(log_vec_2),
  factor_vec_2 = as.numeric(log_vec_2)
)
```

My data set has `r nrow(df_p2)` rows and `r ncol(df_p2)` columns. The mean of variable x is `r mean(pull(df_p2, x))`, the median is `r median(pull(df_p2, x))`, and the standard deviation is `r sd(pull(df_p2, x))`. The proportion of cases for which x + Y > 1 is `r mean(pull(df_p2, log_vec_2))`.

Next I will make a scatterplot of y vs x that color points using the logical, numeric, and factor variables.

```{r}
ggplot(df_p2, aes(x=x, y=y, color = log_vec_2)) + geom_point()
ggsave("scatter_plot.pdf", height = 4, width = 6)
ggplot(df_p2, aes(x=x, y=y, color = num_vec_2)) + geom_point()
ggplot(df_p2, aes(x=x, y=y, color = factor_vec_2)) + geom_point()

```

In the second and third plots the scale of our points is 0-1. Both appear to be highly segmented between 0s and 1s.